Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-24 numerical-mhd-ct.tex
Edited section 4.3 (MHD-CT)

The third solver we describe is an MHD method developed by 
\citet{Collins10} and since a full description and suite of test problems can be
found in that paper, we only describe the method briefly here.

The divergence of the magnetic field, \Divb, is
identically zero due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero.
%\begin{eqnarray}
%\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
%\Divb & =& 0.
%\end{eqnarray}
The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
magnetohydrodynamic (MHD) evolution employs this same vector property
to evolve the magnetic field in a manner that preserves $\Divb=0$.  
The electric field is computed using
the fluxes from the Riemann solver.  The curl of that electric field
is then used to update the magnetic field.  The advantage of this
method is that it preserves \Divb\ to machine precision.  The primary
drawback is increased algorithm complexity.  Note also that since 
only the update of the magnetic field is divergence free, any monopoles 
created by other numerical sources persist.

@@

The dual energy formalism has also been incorporated in two possible ways, one internal
energy, as described in Section \ref{sec.hydro.ppm}, and one that
uses entropy \cite{TVD93, Collins10}.
%The additional equations are
%\begin{eqnarray}
%\end{eqnarray}
%\begin{eqnarray}
%  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
%  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
%  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
%  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
%\end{eqnarray}
%where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
%the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
%alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.

>>>>

The third solver we describe is an MHD method developed by 
\citet{Collins10}.  Since a full description and suite of test problems can be
found in that reference, we only describe the method briefly here.

The divergence of the magnetic field, \Divb, is
identically zero in reality due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero.
%\begin{eqnarray}
%\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
%\Divb & =& 0.
%\end{eqnarray}
The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
magnetohydrodynamic (MHD) evolution employs this same vector property
to evolve the magnetic field in a manner that preserves $\Divb=0$.  
The electric field is computed using
the fluxes from the Riemann solver.  The curl of that electric field
is then used to update the magnetic field.  The advantage of this
method is that it preserves \Divb\ to machine precision.  The primary
drawback is increased algorithm complexity.  Note also that since 
only the update of the magnetic field is divergence free, any monopoles 
created by other numerical sources (such as ill-chosen initial conditions) persist.

@@

The dual energy formalism has also been incorporated in two possible
ways -- one that uses internal
energy, as described in Section \ref{sec.hydro.ppm}, and one that
uses entropy \cite{TVD93, Collins10}.

%The additional equations are
%\begin{eqnarray}
%\end{eqnarray}
%\begin{eqnarray}
%  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
%  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
%  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
%  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
%\end{eqnarray}
%where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
%the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
%alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.
